Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber

### 2014-12-11 pot/LaTeX/skewnormal.tex
adjustment to thresholded observation generation

\section{Spatial skew processes}\label{s:spatialskew}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for non-symmetric data \citep{Genton2004}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

\subsection{Skew-$t$ process} \label{s:skewt}
Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$.
The spatial skew-$t$ process can be written
\begin{align}
  Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
\end{align}
where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is a set of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
It should be noted that the spatial-$t$ process is a special case of the skew-$t$ process when $\lambda = 0$.
A common spatial correlation function is the \Matern with
\begin{align}
  \text{cor}(v(\bs), v(\bt)) = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second time, and $h = || \bs - \bt ||$.

Let $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$ be a set of observations at a finite collection of locations $\bs_1, \ldots, \bs_n$.
After marginalizing over both $z$ and $\sigma$,
\begin{align}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
\end{align}
that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$ where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2013}.

@@

\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
We handle the problem of long-range asymptotic dependence with a random partition model.
As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$to vary by site.
Then, the model becomes
\begin{align} \label{eq:partition}
  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
Let $\bw = (w_1, w_2)$ be the location of a spatial knot.
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
The knots define a random daily partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align*}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align*}
All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
\begin{align}
  z(\bs) &= z_{k} \label{eq:sitez}\\
  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig},
\end{align}
and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.

When incorporating the random daily partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
\begin{align}
  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
\end{align}
where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
Therefore, sites in different subregions are asymptotically independent.
Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
A proof of this is given in Appendix \ref{a:proofsamepartition}.

Figure \ref{fig:simchi} shows $\chi(h)$ for different sample quantiles from simulated datasets where $\bs \in [0, 1] \times [0, 1]$.
These plots show how the partitioning allows for short range extremal dependence while reducing extremal dependence as the distance between sites increases.

\begin{figure}
  \includegraphics[width=\linewidth]{plots/sim-chi.pdf}
  \caption{$\chi(h)$ for simulated datasets. The dashed line indicates the $\chi(h)$ we would expect under asymptotic independence.}
  \label{fig:simchi}
\end{figure}

\subsection{Extension to space-time data} \label{s:temporal}
When using daily measurements, the assumption of temporal independence is inappropriate.
There are several places where temporal dependence could be incorporated in the model, including the residual $v_t(\bs)$.
However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
Let
\begin{align*}
  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
\end{align*}
where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
As in section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
\begin{align}
  z_t(\bs) = z_{tk}\\
  \sigma_{t}(\bs) = \sigma_{tk}.
\end{align}

We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align}
  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(s_i))}{ \text{range}(s_i) } \right], \quad i = 1, 2
\end{align}
where $\Phi$ is a univariate standard normal density function.
Then $\bw^*_k \in \calR^2$.
We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align}
where IG is the distribution function for an IG$(a, b)$ random variable.
The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z_{tk} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
We restrict $\phi_z$ to be positive because in the skew-$t$ distribution, we use $|z_t(\bs)|$.
These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2 \sim \text{IG}(a, b)$.
For each day, the model is identical to the spatial-only model in Section \ref{s:part} by construction.

\subsection{Hierarchical model}\label{s:hier}
Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
We model this with a Bayesian hierarchical model as follows.
Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \ref{s:part}.

Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
   \lambda &\sim N(0, 10)\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
\end{align}
where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.

>>>>

\section{Spatial skew processes}\label{s:spatialskew}
Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Genton2004}.
A brief review of the additive process by which a skew-$t$ process is created is given here.

\subsection{Skew-$t$ process} \label{s:skewt}
Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$.
The spatial skew-$t$ process can be written
\begin{align}
  Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
\end{align}
where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is a set of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
A common spatial correlation function is the \Matern with
\begin{align}
  \text{cor}(v(\bs), v(\bt)) = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
\end{align}
where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.

Let $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$ be a set of observations at a finite collection of locations $\bs_1, \ldots, \bs_n$.
After marginalizing over both $z$ and $\sigma$,
\begin{align}
  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
\end{align}
that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2013}.

@@

\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
We handle the problem of long-range asymptotic dependence with a random partition model.
As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
Then, the model becomes
\begin{align} \label{eq:partition}
  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
Let $\bw = (w_1, w_2)$ be the location of a spatial knot.
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
The knots define a random daily partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{eq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
\begin{align}
  z(\bs) &= z_{k} \label{eq:sitez}\\
  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig},
\end{align}
and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.

When incorporating the random daily partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
\begin{align}
  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
\end{align}
where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
Therefore, sites in different subregions are asymptotically independent.
Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$.
A proof of this is given in Appendix \ref{a:proofsamepartition}.

Figure \ref{fig:simchi} shows $\chi(h)$ for varying data settings when $\bs \in [0, 1] \times [0, 1]$.
The plot clearly shows that without partitioning, in a skew-$t$ process, all sites remain asymptotically dependent.
However, upon incorporating partitioning into the model, the asymptotic dependence is only shared by sites that are close to one another.

\begin{figure}
  %\includegraphics[width=\linewidth]{plots/sim-chi.pdf}
  \caption{$\chi(h)$ for different settings. I need to replace the plot here.}
  \label{fig:simchi}
\end{figure}

\subsection{Extension to space-time data} \label{s:temporal}
When using daily measurements, the assumption of temporal independence is inappropriate.
There are several places where temporal dependence could be incorporated in the model, including the residual $v_t(\bs)$.
However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
Let
\begin{align} \label{eq:spatiotemp}
  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
\end{align}
where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
As in section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
\begin{align}
  z_t(\bs) = z_{tk}\\
  \sigma_{t}(\bs) = \sigma_{tk}.
\end{align}

We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
Let
\begin{align}
  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(\bs_i))}{ \text{range}(\bs_i) } \right], \quad i = 1, 2
\end{align}
where $\Phi$ is a univariate standard normal density function, and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
Then $\bw^*_{tk} \in \calR^2$.
We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
Let
\begin{align}
  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
\end{align}
where IG is the distribution function for an IG$(a, b)$ random variable.
The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
\begin{align}
  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
  z_{tk} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
\end{align}
where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
We restrict $\phi_z$ to be positive because in the model, we assume $Y(\bs)$ is a function of $|z_t(\bs)|$.
These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2_{tk} \sim \text{IG}(a, b)$.
For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.

\subsection{Hierarchical model}\label{s:hier}
Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
We model this with a Bayesian hierarchical model as follows.
Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).

Then
\begin{align}
   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
   \lambda &\sim N(0, 10)\\
   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
\end{align}
where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
